Watch Out for the Bully! Study of Interference Between Jobs Running Concurrently on Dragonfly Network

### 2016-03-15 draft.tex
finish 1st version

The high-radix, low-diameter dragonfly topology network will be the choice of many computing facilities for building their next generation supercomputers. Preliminary study shows that random job allocation with adaptive routing would be the rule of thumb to utilize such powerful network. However, in this work, we made the observation that although random job allocation with adaptive routing can alleviate the hotspot and make network traffic load-balanced for the workload, it can not guarantee the best performance for each single job. The performance improvement of communication intensive jobs comes with sacrifice of those less intensive jobs. We refer this phenomenon as ``bully" between concurrently running jobs. In this paper, we analyze the detailed reason for the happening of this ``bully" from the network level. Our study is based on a sophisticated, high-fidelity discrete even-driven simulation tool. We believe the observations and analysis in this work will make the chosen of job allocation and routing strategies more cautiously in future.

@@

@@

The low latency and high bandwidth interconnect network plays a critical role in improving HPC system performance. As the computing ability of HPC system continue grows due to many-core nodes with accelerators, the network is now becoming a scare resources. HPC community start to explore the possible alternative to replace the commonly used Torus network.

The building high-radix, low-diameter dragonfly topology can lower the cost of interconnect network, improve network bandwidth, reduce the packets traverse latency. This make the dragonfly network an very promising choice for building multi-Petaflop/s supercomputers\cite{dally-dragonfly}. The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

@@

This is paper, we study the interference between concurrently running jobs on dragonfly network with different job placement and routing schemes. Being unorthodox to most existing work that use synthetic workload, our work is based on application traces that collected from production system. The advantage of using real application traces is quite obvious. The traces collected from production system can capture application's communication behavior in detail that usually omitted in synthetic workload for the purpose of simplicity.

@@

We made observation from our extensive study that, when dragonfly network enabled with random job placement and adaptive routing scheme in order to reach hot-spot free and load balance, the performance of particular job(s) in the workload would suffer degradation. The communication intensive jobs in the workload can benefit from random placement and adaptive routing, with great performance improvement. Unfortunately, those less intensive jobs are the sacrifices for this improvement, their performance will be greatly reduced.

@@

    \item Our real trace-based analysis about job interference on dragonfly network, shows that random placement with adaptive routing can not guarantee the performance of every job in the workload. The hotspot-free and load-balanced reached by using random placement with adaptive routing come with the cost of communication-less-intensive jobs performance degradation.

    \item We identify the ``bully" behavior happening between jobs running concurrently on dragonfly network with random placement and adaptive routing. Random placement and adaptive routing redirect the traffic of communication intensive jobs to the routers through which less intensive jobs' traffic used to go, embezzle the network resources, slow down the communication of those less intensive jobs.

    \item We found that contiguous placement and minimal routing can guarantee the consistent performance of each job running on dragonfly network, while may leads to local hotspot and severe network congestion. Actually, contiguous placement and minimal routing are preferred by communication-less-intensive jobs, since it will prevent them from being bullied by other communication-intensive jobs.

@@

To the best of our knowledge, using application traces from production system for the study of job interference on dragonfly network as not been reported so far. We believe the observations and analysis presented in this paper can be used by computing facilities to choose job placement and routing strategies wisely for their future dragonfly interconnected system. The conclusion from this paper would be useful for batch schedulers in dragonfly system to make better scheduling decisions for workload consists of jobs that have various communication behavior.

@@

\subsection{Dragonfly Network Overview}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by All-to-All links. In each group, there could be ``a" number of routers connected via All-to-All local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal link. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other router in different groups. Thus, the radix of each router is $k = a+h+p-1$. The implementation of such dragonfly network could be different between different computing facilities. The choosing of proper $a$, $h$, $p$ could be involving many factors, such as system scale, building cost and workload requirements.

For the purpose of loading balance, it is recommended that the proper dragonfly configuration is $a=2p=2h$\cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g = $.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. Illustration of part of dragonfly network. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job\_1 requires for 6 nodes, get the allocation by Random placement. Job\_2 requires for 8 nodes, get the allocation by Contiguous placement. }
  \label{fig: dragonfly overview}
\end{figure}

@@

Job placement refers to the scheme used by batch scheduler to allocate the required number of nodes by the job for execution. As studied by several researchers, the job placement could be an important factor that impact application performance \cite{hoefler-hpdc14} \cite{bhatele-sc11} \cite{jain-sc14}. In this work, we study the following two placement schemes:

\textbf{Contiguous Placement:} In this scheme, the nodes will first be indexed consecutively in the increasing oder of their group id and router id. The job is allocated by cutting chunks of nodes from the sequentially ordered list, as illustrated in Figure\ref{fig: dragonfly overview} as blue nodes. The nodes been allocated to each job will be adjacent to each other, usually they are attached to the same router, in the same group. Contiguous placement confines the traffic of application into the local area, which may leads to local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this scheme, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes allocated to each job would be attached to different routers in different groups, as illustrated in Figure\ref{fig: dragonfly overview} as red nodes. Some routers may be shared by different applications when using random placement. Random placement can spreading the traffic of application uniformly across the whole system, thus avoid the possible hot-spots and reach load balance.

@@

The implemented routing schemes in CODES dragonfly network model are based on the routing algorithm proposed by Kim et al. \cite{dally-dragonfly}. Two routing schemes studied in this work are \emph{Minimal} and \emph{Adaptive}.

\textbf{Minimal:} The minimal routing sends a message from source router to destination router along the shortest path. When there are multiple shortest paths, the message is evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may leads to congestion along the shortest path.

\textbf{Adaptive:} In this scheme, the path a message takes will be adaptively chosen between shortest and non-shortest path, depends on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message will first take the shortest path from source to that intermediate router, then it will take the shortest path between the intermediate router to the destination. Since adaptive scheme chooses routing path adaptively based on the congestion situation, it can void local congestion and reduce the possibility of hot-spots.

\textbf{Progressive Adaptive:} \NOTE{say something about prog-adaptive routing.} \cite{won-prog-adaptive}.

@@

\subsection{Simulation Tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability\cite{codes}.  CODES support dragonfly network with high fidelity packet-level simulation\cite{codes-dragonfly}. It has this network workload component that is capable of taking real MPI application traces generated by SST DUMPI\cite{sst} for trace-driven simulations.

@@

We choose three representative applications from DOE Design Forward Project\cite{designforwardwebpage}. We believe these applications can be representative of a wide array of applications running on leadership-class supercomputers. Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.

AMG: The Algebraic MultiGrid Solver, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}. The traces of AMG we chosen are comprised of 216 MPI ranks.

MultiGrid: massively parallel block-structured adaptive mesh refinement (AMR) codes a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes\cite{boxlib}. It is widely used for structured grid physics packages. The traces of MultiGrid we chosen are comprised of 125 MPI ranks.

CrystalRouter: the extracted communication kernel of the full application Nek5000\cite{nek5000}, which is a spectral element CFD application developed at Argonne National Laboratory\cite{crystalrouter}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores,and spatiotemporal chaos. The traces of CrystalRouter we chosen are comprised of 100 MPI ranks.

@@

@@

There are 6 placement and routing scheme combinations, namely contiguous placement coupled with minimal routing, contiguous placement coupled with adaptive routing, contiguous placement coupled with progressive adaptive routing random placement coupled with minimal routing, random placement coupled with adaptive routing, random placement coupled with progressive adaptive routing.  In this work, we enable the applications running on dragonfly network under different placement and routing configurations, and scrutinize the application's performance and network behavior in different running scenarios.

\begin{table}[ht]
\begin{center}
\caption{Job placement and routing configurations} 
\label{tab: placement routing configs}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Routing Schemes} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
Placement schemes & Minimal & Adaptive & Progressive adp\\ % Column names row
\midrule % In-table horizontal line
Contiguous  &     &      &     \\ % Content row 1
\midrule
Random  &     &      &     \\

@@

The results of random placement shown in the following sections are collected from extensive runs of different random placements, which can eliminate the possibility of variations.

\section{Parallel Workload \Rmnum{1 } Study}
\label{sec: workload 1 overview}

The first parallel workload analyzed in our experiment consists of three applications, namely AMG , MultiGrid (MG) and CrystalRouter(CR). The details of this workload shown in Table \ref{tab: parallel workload-1}.

\begin{table}[ht]
\begin{center}
\caption{Parallel Workload \Rmnum{1 }} 
\label{tab: parallel workload-1}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Routing Schemes} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
Application & num. rank & Data per Rank & Total Data\\ % Column names row
\midrule % In-table horizontal line
AMG  &    216 &   0.6MB   &     \\ % Content row 1
\midrule
MultiGrid  &    125 &   5MB   &     \\ 
\midrule
CrystalRouter  &   100  &  35MB    &     \\

Figure \ref{fig:3app-data-amount} shows the data transfer amount of each MPI rank of three application. Based on this figure, we could say that CrystalRouter is the most communication-intensive and AMG is the lest communication-intensive.

@@

In our experiments, the workload running on the dragonfly network under 6 different job placement and routine scheme configurations as listed in Table \ref{tab: placement routing configs}. We focus on the analysis of network status while the workload is running, as well as the performance of each individual application. We collected the following result data from the experiments:
\begin{itemize}
    \item router terminal link traffic and busy time.
    \item router local channel traffic and busy time.
    \item router global channel traffic and busy time.
    \item The time application spend on communication. 
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{wkld/wkld-commtime}
  \caption{Time spent on communication by Workload I when 6 different placement and routing scheme configurations are used. Random is better than contiguous in terms of improving communication efficiency.Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
  \label{fig:wkld-commtime}
\end{figure}

@@

\NOTE{The take-home message} from Figure \ref{fig:wkld-network-stime} and \ref{fig:wkld-network-traffic}.

@@

@@

\section{Parallel Workload \Rmnum{2 } Study}

@@

@@

\TODO{Explain} Figure \ref{fig:syn-wkld-commtime}.

The communication time of each application shown in Figure \ref{fig:syn-apps-commtime}. The ``bully" become the ``bullyee". The less communication-intensive applications in the workload this time are  MultiGrid and CrystalRouter.  When random placement coupled with (progressive) adaptive routing is used, MultiGrid and CrystalRouter suffer prolonged communication time, shown in Figure \ref{fig:syn-mg-commtime}, \ref{fig:syn-cr-commtime}. sAMG on the other hand benefit from random placement, shown in Figure \ref{fig:syn-samg-commtime }.

@@

@@

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

@@

The authors would like to thank...

>>>>

The high-radix, low-diameter dragonfly topology network will be the choice of many computing facilities for building their next generation supercomputers. Preliminary study shows that random placement with adaptive routing should be the rule of thumb to utilize such powerful network, since random placement can uniformly distribut traffic over the network, and adaptive routing can alleviate congestion. However, in this paper, we made the observation that although random placement coupled with adaptive routing can reach load-balanced over the network when multiple jobs running concurrently, it can not guarantee the best performance for each single job. The performance improvement of communication intensive jobs comes with sacrifice of those less intensive jobs. We refer this phenomenon as ``bully" between concurrently running jobs. We analyze the detailed reason behind this ``bully" behavior between jobs from the underneath network level. Our study is based on a sophisticated, high-fidelity discrete even-driven simulation tool. We believe the observations and analysis in this work will make the chosen of job placement and routing strategies on dragonfly network more cautiously in future.

@@

% \keywords{Dragonfly, Job placement, Routing, Interference}

@@

The low latency and high bandwidth interconnect network plays a critical role in improving HPC system performance. As the computing ability continue grows due to many-core nodes with accelerators, the network is now becoming a scare resources. HPC community start to explore the possible alternative to replace the commonly used Torus network.

The high-radix, low-diameter dragonfly topology can lower the cost of interconnect network, improve network bandwidth, reduce the packets traverse latency. This make the dragonfly network an very promising choice for building multi-Petaflop/s supercomputers\cite{dally-dragonfly}. The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

@@

This is paper, we study the interference between concurrently running jobs on dragonfly network with different job placement and routing schemes. Being unorthodox to most existing work that use synthetic workload, parallel scientific application traces that collected from production system are used in our study. The advantage of using real application traces is quite obvious, these traces can capture application's communication behavior in detail that usually omitted in synthetic workload for the purpose of simplicity. Applicatoins usually are distinctive from each other in terms of  data transfer amount, communication pattern, operation dependency etc. which are impossible to reproduce in synthetic workload.

@@

We made observation from our extensive study that, when dragonfly network is configured with random placement and adaptive routing scheme in order to reach hot-spots free and load balance, the performance of particular job(s) in the workload would suffer degradation. The communication intensive jobs in the workload can benefit from random placement and adaptive routing, with great performance improvement. Unfortunately, those less intensive jobs are the sacrifices for this improvement, their performance will be greatly reduced. We identify this phenomenon as ``bully" between concurrently running jobs.

@@

    \item Our real trace-based analysis about job interference on dragonfly network, shows that random placement with adaptive routing can not guarantee the performance of every job in the workload. The hot-spots free and load-balanced reached by using random placement with adaptive routing come with the cost of less communication-intensive jobs performance degradation.

    \item We identify the ``bully" behavior happening between concurrently running jobs on dragonfly network when random placement and adaptive routing in use. Random placement and adaptive routing redirect the traffic of communication intensive jobs to the routers through which less intensive jobs' traffic used to go, embezzle the network resource such as local and global channel bandwidth, slow down the communication of those less intensive jobs.

    \item We found that contiguous placement and minimal routing can guarantee the consistent performance of each job running on dragonfly network, while may result in hot-spots and local network congestion. Actually, contiguous placement and minimal routing are preferred by communication-less-intensive jobs, since it will prevent them from being ``bullied" by other communication-intensive jobs.

@@

To the best of our knowledge, using parallel application traces from production system for the study of job interference on dragonfly network has not been reported so far. We believe the observations and analysis presented in this paper can be used by computing facilities to choose job placement and routing strategies wisely for their future dragonfly interconnected system. The conclusion from this paper would be useful for workload manager in dragonfly system to make better scheduling decisions for jobs with distinctive communication behavior.

@@

\subsection{Dragonfly Network Overview}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by All-to-All links. In each group, there could be ``a" number of routers connected via All-to-All local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. The implementation of such dragonfly network could be different between different computing facilities. The choosing of proper $a$, $h$, $p$ could be involving many factors, such as system scale, building cost and workload requirements.

For the purpose of loading balance, it is recommended that the proper dragonfly configuration should be like $a=2p=2h$\cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. Figure \ref{fig:dragonfly overview} illustrated an implementation of dragonfly network follows such configuration. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. Illustration of part of dragonfly network. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job\_1 requires for 6 nodes, get the allocation by Random placement. Job\_2 requires for 8 nodes, get the allocation by Contiguous placement. }
  \label{fig:dragonfly overview}
\end{figure}

@@

Job placement refers to the scheme used by batch scheduler to allocate the required number of nodes by each job for execution. As studied by several researchers, the job placement could be an important factor that impact application performance \cite{hoefler-hpdc14} \cite{bhatele-sc11} \cite{jain-sc14}. In this work, we study the following two placement schemes that are most likely to be deployed in the future dragonfly system.

\textbf{Contiguous Placement:} In this scheme, the nodes will first be indexed consecutively in an increasing order of their group id and router id. The schedul assign each job required number of nodes by cutting chunks of nodes from the sequentially ordered list, as illustrated in Figure \ref{fig: dragonfly overview} as blue nodes assigned to job\_2. The nodes been allocated to the same job will be adjacent to each other, usually they are attached to the same router or in the same group, depends on the size of the job. Contiguous placement confines the traffic of application into the local area, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this scheme, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes allocated to each job would be attached to different routers in different groups, as illustrated in Figure \ref{fig: dragonfly overview} the red nodes assigned to job\_1. Some routers may be shared by different applications when using random placement. Random placement can spreading the traffic of application uniformly across the whole system, thus avoid the possible local congestion and reach load balance.

@@

The implemented routing schemes in CODES dragonfly network model are based on the routing algorithm proposed by Kim et al. \cite{dally-dragonfly} and Jiang et al. \cite{jiang}. Three routing schemes are studied in this work are \emph{Minimal}, \emph{Adaptive} and \emph{Progressive Adaptive}.

\textbf{Minimal:} When the network deploy minimal routing, a message from source router will take the the shortest path to the destination router. When there are multiple shortest paths between the rource and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this scheme, the path a message takes will be adaptively chosen between shortest and non-shortest path, depends on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message will first take the shortest path from source to that intermediate router, then it will take the shortest path from that intermediate router to the destination. Since adaptive scheme chooses routing path adaptively based on the congestion situation, it can void local congestion and reduce the possibility of hot-spots.

\textbf{Progressive Adaptive:} In this scheme, the decision to route minimally at each hop in the source group will be re-evaluated and any decision to route non-minimally at source router or at a subsequent hop is permant and is not re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

\subsection{Simulation Tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability\cite{codes}.  CODES supports dragonfly network with packet-level high-fidelity simulation\cite{codes-dragonfly}. It has this network workload component that is capable of taking real MPI application traces generated by SST DUMPI\cite{sst} for trace-driven simulations.

@@

We choose three representative parallel applications from DOE Design Forward Project\cite{designforwardwebpage}. We believe these applications can be representative of a wide array of applications running on leadership-class supercomputers. Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.

AMG: The Algebraic MultiGrid Solver, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}. 
% The traces of AMG we chosen are comprised of 216 MPI ranks.

MultiGrid:  geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes\cite{boxlib}. It is widely used for structured grid physics packages. 
% The traces of MultiGrid we chosen are comprised of 125 MPI ranks.

CrystalRouter: the extracted communication kernel of the full application Nek5000\cite{nek5000}, which is a spectral element CFD application developed at Argonne National Laboratory\cite{crystalrouter}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores,and spatiotemporal chaos. 
% The traces of CrystalRouter we chosen are comprised of 100 MPI ranks.

@@

\begin{table}[ht]
\begin{center}
\caption{The notation for different placement \& routing configurations} 
\label{tab: placement routing configs}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Routing} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
Placement  & Minimal & Adaptive & Progressive adaptive\\ % Column names row
\midrule % In-table horizontal line
Contiguous  &  CM   &   CA   &  CPA   \\ % Content row 1
\midrule
Random  &   RM  &   RA   &  RPA   \\

@@

We focus on the study of two placement schemes and three routing schemes, so there will be 6 different placement\&routing configurations, as listed in Table \ref{tab: placement routing configs}. In this work, we enable the applications running on dragonfly network under different placement and routing configurations, and scrutinize the application's performance and network behavior in different running scenarios.

In our experiments, the workload running on the dragonfly network under 6 different placement and routing configurations as listed in Table \ref{tab: placement routing configs}. We focus on the analysis of network status while the workload is running, as well as the performance of each individual application. We collected the following result data from the experiments:

\begin{itemize}
    \item router terminal link traffic and busy time.
    \item router local channel traffic and busy time.
    \item router global channel traffic and busy time.
    \item The time spend on communication for each MPI rank. 
\end{itemize}

The results of random placement shown in the following experiments are collected from extensive runs of different random placements, which can eliminate the possibility of variations.

\section{Parallel Workload \Rmnum{1 }}
\label{sec: workload 1 overview}

The first parallel workload analyzed in our experiment consists of three applications presented in section\ref{sec: application traces} . The details of this workload shown in Table \ref{tab: parallel workload-1}.

\begin{table}[ht]
\begin{center}
\caption{Details of Parallel Workload \Rmnum{1 }.} 
\label{tab: parallel workload-1}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Application Details} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
App Name & num. rank & Data per Rank & Total Data\\ % Column names row
\midrule % In-table horizontal line
AMG  &    216 &   0.6MB   &     130MB\\ % Content row 1
\midrule
MultiGrid  &    125 &   5MB   &     625MB\\ 
\midrule
CrystalRouter  &   100  &  35MB    &     3500MB\\

@@

@@

Figure \ref{fig:3app-data-amount} shows the data transfer amount of each MPI rank of three application. Based on this figure, we can tell that CrystalRouter is the most communication-intensive and AMG is the lest communication-intensive.

\subsection{Workload Analysis}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 0.48\textwidth ]{wkld/wkld-commtime}
  \caption{Time spent on communication by Workload I when 6 different placement and routing configurations are used. Random is better than contiguous in terms of improving communication efficiency.Communication time can be further reduced when (progressive) adaptive routing is used under both placement schemes. }
  \label{fig:wkld-commtime}
\end{figure}

The communication time spent by parallel workload \Rmnum{1} under different placement and routing configuration shown in Figure \ref{fig:wkld-commtime}. The contiguous placement coupled with minimal routing results in highest communication time. The random placement coupled with minimal routing is the most efficient. Progressive adaptive routing is basically the same as adaptive when coupled with both placement schemes.

@@

% \NOTE{The take-home message} from Figure \ref{fig:wkld-network-stime} and \ref{fig:wkld-network-traffic}.

@@

\subsection{Individual Application Analysis}

@@

\section{Parallel Workload \Rmnum{2 }}

\begin{table}[ht]
\begin{center}
\caption{Details of Parallel Workload \Rmnum{2 }.} 
\label{tab: parallel workload-2}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{3}{c}{Application Details} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-4}
App Name & num. rank & Data per Rank & Total Data\\ % Column names row
\midrule % In-table horizontal line
sAMG  &    216 &   60MB   &     13000MB\\ % Content row 1
\midrule
MultiGrid  &    125 &   5MB   &     625MB\\ 
\midrule
CrystalRouter  &   100  &  35MB    &     3500MB\\

@@

\subsection{Workload Analysis}

@@

The communication time spent by parallel workload \Rmnum{2} under different placement and routing configuration shown in Figure \ref{fig:syn-wkld-commtime}. As the data transfer amount increased in the workload(due to sAMG), the time spent on communication under each configuration also increases. But the performance of each configuration still keep the same trend as parallwl workload \Rmnum{1}. The contiguous placement coupled with minimal routing results in highest communication time. The random placement coupled with minimal routing is the most efficient. Progressive adaptive routing is better than adaptive when random placement is used. And they have comparable results when contiguous placement is used.

\\ \TODO{Explain Figure }  \ref{fig:synwkld-network-traffic} \ref{fig:synwkld-network-stime}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:synwkld-global-channel-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:synwkld-local-channel-traffic}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:synwkld-terminal-link-traffic}
    \end{subfigure}%
  \caption{Network Traffic. \textbf{More explanation to be added for this figure}}
   \label{fig:synwkld-network-traffic}
\end{figure*}

\begin{figure*}[t!]
    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/gc-btime}
        \caption{Global Channel Busy Time}
        \label{fig:synwkld-global-channel-stime}
    \end{subfigure}%
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/lc-btime}
        \caption{Local Channel Busy Time}
        \label{fig:synwkld-local-channel-stime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/tl-btime}
        \caption{Terminal Link Busy Time}
        \label{fig:synwkld-terminal-link-stime}
    \end{subfigure}%
   \caption{Network Channel Busy Time. \textbf{More explanation to be added for this figure}}
   \label{fig:synwkld-network-stime}
\end{figure*}

\subsection{Individual Application Analysis}

The communication time of each application shown in Figure \ref{fig:syn-apps-commtime}. The ``bully" become the ``bullyee". The less communication-intensive applications in the workload this time are  MultiGrid and CrystalRouter.  When random placement coupled with (progressive) adaptive routing is used, MultiGrid and CrystalRouter suffer prolonged communication time, shown in Figure \ref{fig:syn-mg-commtime}, \ref{fig:syn-cr-commtime}. sAMG on the other hand benefit from random placement, shown in Figure \ref{fig:syn-samg-commtime}.

@@

\section{Summary}

@@

@@

The work at Illinois Institute of Technology is supported in part by U.S. National Science Foundation grants CNS-1320125 and CCF-1422009. This work is also supported by the U.S. Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
